---
title: "How Prevalent is Archaeological Garden Hunting?"
author: "Jonathan Dombrosky, Caitlin Ainsworth, Abigail Judkins, Jana Meyer, Michael A. Adler, and Emily Lena Jones"
date: "1/24/2022"
output:
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, dpi = 400)
```

# Libraries
```{r, message = FALSE, warning = FALSE}
library(SIBER)
library(tidyverse)
library(cowplot)
library(broom)
library(car)
```

# Figure 1
Plant genera reported by Scribner and Krysel (1982) were assigned a
photosynthetic pathway from the following sources:
Basinger and Robertson (1997), Bruhl and Wilson (2007), Danneberger (1999),
Giussani et al. (2001), Kocacinar and Sage (2003), Nelson (2012),
Osborne et al. (2014), and Syvertsen et al. (1976).
```{r, warning = FALSE, fig.width=8, fig.height=4}
Table_2 <- read.csv("Scribner and Krysl_Table 2/Table 2.csv", header = TRUE)

Pathway <- Table_2 %>%
  select(Environmental.Context, Photosynthetic.pathway, DF....) %>%
  group_by(Environmental.Context, Photosynthetic.pathway) %>%
  summarize(sum_DF = sum(DF....), .groups = "keep")

labels <- c("Agricultural Playa Basins" = "Agricultural Context",
            "Playa Basins" = "Non-Agricultural Context")

Pathway %>%
ggplot(aes(x = reorder(Photosynthetic.pathway, -sum_DF), y = sum_DF)) +
  geom_bar(stat = "identity", size = 0.75, alpha = 0.5, color = "#4d4d4d") +
  facet_wrap(~ Environmental.Context, ncol = 2, 
             labeller = labeller(Environmental.Context = labels)) +
  scale_x_discrete(labels = parse(text = c("C[4]", "C[3]", "C[3]/C[4]"))) +
  theme_classic() +
  theme(legend.position="none",
        strip.background = element_blank(),
        strip.text.x = element_text(color = "#4d4d4d", size = 12, 
                                    face = "bold"),
        axis.line = element_line(color = "#4d4d4d", size = 0.75),
        axis.text.x = element_text(color = "#4d4d4d", size = 10),
        axis.text.y = element_text(color = "#4d4d4d", size = 10),
        axis.title.x = element_text(color = "#4d4d4d", size = 12, 
                                    face = "bold"),
        axis.title.y = element_text(color = "#4d4d4d", size = 12, 
                                    face = "bold"),
        axis.ticks.x = element_line(color = "#4d4d4d", size = 0.75),
        axis.ticks.y = element_line(color = "#4d4d4d", size = 0.75)) +
        labs(x = "Photosynthetic Pathway", y = "% Diet")
```

# Isotope Data
Bone from the `archaeo` dataset was demineralized in 0.5 N hydrochloric acid, 
lipids were removed using 2:1 chloroform:methanol, and the resulting collagen 
pseudomorph was freeze-dried overnight. Between 0.5 and 0.6 mg of collagen was 
weighed out for the analysis of δ^13^C and δ^15^N. All seeds from the `seeds` 
dataset were purchased from Native seeds/SEARCH. Between 5.0 and 6.0 mg of 
ground corn and 2.0 and 2.5 mg of ground bean/squash was weighed out for the 
analysis of δ^13^C and δ^15^N. δ^13^C and δ^15^N were measured at the 
University of New Mexico Center for Stable Isotopes (UNM CSI, Albuquerque, NM) 
on a Thermo Scientific Delta V isotope ratio mass spectrometer (IRMS) with a 
dual inlet and Conflo IV interface coupled to a Costech 4010 elemental analyzer
(EA). Stable isotope values are reported as parts per mil (‰).

The `humans` and `turkeys` isotope values come from the following sources: 
Chisholm and Matson (1994), Coltrain et al. (2007), Conrad et al. (2016), 
Jones et al. (2016), Kellner et al. (2010), Kennett et al. (2017), 
Martin (1999), McCaffery et al. (2014), and Rawlings and Driver (2010).
```{r}
archaeo <- read.csv("archaeological.csv", header = TRUE)
seeds <- read.csv("modern seeds.csv", header = TRUE)
turkeys <- read.csv("turkeys.csv", header = TRUE) 
humans <- read.csv("humans.csv", header = TRUE)
```

# Assessing Collagen Purity
Boxplots of C:N~atomic~ values of archaeological leporid collagen per site.
The blue box represents the acceptable range of collagen purity (2.9-3.6) 
reported by Ambrose (1990). 
```{r, fig.width = 7, fig.height = 5}
archaeo %>%
  mutate(CNatomic = CN * (14/12)) %>%
  ggplot(mapping = aes(y = CNatomic, x = Site.Name, group = Site.Name)) +
  geom_boxplot(color = "#4d4d4d", size = 0.75) +
  labs(y = expression("C:N"[atomic]), x = "Archaeological Site") +
  theme_classic() +
  theme(legend.position="none",
        axis.line = element_line(color = "#4d4d4d", size = 1),
        axis.text.x = element_text(color = "#4d4d4d", size = 12),
        axis.text.y = element_text(color = "#4d4d4d", size = 12),
        axis.title.x = element_text(color = "#4d4d4d", size = 14),
        axis.title.y = element_text(color = "#4d4d4d", size = 14),
        axis.ticks.x = element_line(color = "#4d4d4d", size = 1),
        axis.ticks.y = element_line(color = "#4d4d4d", size = 1)) +
    scale_y_continuous(limits=c(2.9, 3.6)) +
  annotate(geom = "rect", xmin = -Inf, xmax = Inf, ymin = 2.9, ymax = 3.6, 
           color = "#5da5d8", fill =  "#5da5d8", alpha = 0.5, size = 1)
```

# Data Wrangling
A ^13^C Suess correction of 2.0‰ was applied to the modern seed data 
(Dombrosky 2020). 
```{r}
seeds <- seeds %>%
  mutate(d13Csuess = d13C + 2)

archaeo_SIBER <- archaeo %>%
  unite(group, Site.Name, Genus, sep = " ") %>%
  select(group, d13C, d15N)

seeds_SIBER <- seeds %>%
  select(Comparative.Group, d13Csuess, d15N) %>%
  rename(group = Comparative.Group,
         d13C = d13Csuess)

turkeys_SIBER <- turkeys %>%
  mutate(animal = "Turkey") %>%
  unite(group, Diet.Type, animal, sep = " ") %>%
  select(group, d13C, d15N)
  

humans_SIBER <-  humans %>%
  mutate(group = "Humans") %>%
  select(group, d13C, d15N)

SIBER_data <- rbind(archaeo_SIBER, seeds_SIBER, turkeys_SIBER, humans_SIBER)
```

# Figure 2
## Sand Canyon Pueblo Figures
```{r, fig.width = 7, fig.height = 5}
sand_label_df <- data.frame(
  group = c("Sand Canyon Pueblo Lepus", "Sand Canyon Pueblo Sylvilagus",
            "Free-ranging Turkey", "Humans", "Maize-fed Turkey", "Bean",
            "Squash", "Corn"),
  label = c("Sand Canyon Pueblo\nJackrabbbits",
            "Sand Canyon Pueblo\nCottontails",
            "Free-Ranging\nTurkeys", "Humans", "Maize-Fed Turkeys", "Beans",
            "Squash", "Maize"),
  d13C = c(-17, -20, -13.25, -5.8, -8, -25.25, -24, -9.75),
  d15N = c(-0.75, 11, 2.5, 12.2, 4.1, -2, 13.25, -1),
  hjust = c(0.5, 0.5, 0.5, 0, 0, 1, 1, 1),
  vjust = c(1, 0, 0, 1, 1, 1, 0, 1))

sand_label_df$group <- factor(sand_label_df$group, 
                          levels = c("Sand Canyon Pueblo Lepus",
                                     "Sand Canyon Pueblo Sylvilagus",
                                     "Free-ranging Turkey", "Humans", 
                                     "Maize-fed Turkey", "Bean", "Squash",
                                     "Corn"))

sand_plot <- SIBER_data %>%
  filter(group %in% c("Sand Canyon Pueblo Lepus", 
                      "Sand Canyon Pueblo Sylvilagus",
                      "Bean", "Corn", "Squash", "Free-ranging Turkey",
                      "Maize-fed Turkey", "Humans"))
  
sand_plot$group <- factor(sand_plot$group, 
                          levels = c("Sand Canyon Pueblo Lepus",
                                     "Sand Canyon Pueblo Sylvilagus",
                                     "Free-ranging Turkey", "Humans", 
                                     "Maize-fed Turkey", "Bean", "Squash",
                                     "Corn")) 
  
sand_p1 <- ggplot(sand_plot, aes(x = d13C, y = d15N)) +
  geom_point(aes(fill = group, color = group), stroke = 1, size = 4, 
             alpha = 0.5, shape = 21) +
  geom_point(aes(color = group), fill = NA, stroke = 1, size = 4, 
             shape = 21) +
  ylab(expression(paste(delta^{15}, "N (\u2030)"))) +
  xlab(expression(paste(delta^{13}, "C (\u2030)"))) + 
  theme_classic() +
  theme(legend.position = "none",
        axis.line = element_line(color = "#4d4d4d", size = 1),
        axis.text.x = element_text(color = "#4d4d4d", size = 12),
        axis.text.y = element_text(color = "#4d4d4d", size = 12),
        axis.title.x = element_text(color = "#4d4d4d", size = 14, 
                                    face = "bold"),
        axis.title.y = element_text(color = "#4d4d4d", size = 14, 
                                    face = "bold"),
        axis.ticks.x = element_line(color = "#4d4d4d", size = 1),
        axis.ticks.y = element_line(color = "#4d4d4d", size = 1)) + 
  scale_color_viridis_d() +
  stat_ellipse(aes(group = interaction(group), color = group, fill = group), 
               alpha = 0.25, size = 0.75, linetype = 3, level = 0.95, 
               type = "t", geom = "polygon") + 
  geom_text(data = sand_label_df,aes(x = d13C, y = d15N, 
                                label = label, color = group, hjust = hjust,
                                vjust = vjust),
            size = 9/.pt, fontface = "bold") +
  scale_fill_viridis_d() +
  scale_x_continuous(limits=c(-27.5, -4), 
                     breaks = c(-25, -20, -15, -10, -5)) +
  scale_y_continuous(limits=c(-3, 15))

sand_p1

sand_p2 <- ggplot(sand_plot, aes(x = d13C, y = d15N)) +
  ylab(expression(paste(delta^{15}, "N (\u2030)"))) +
  xlab(expression(paste(delta^{13}, "C (\u2030)"))) + 
  theme_classic() +
  theme(legend.position = "none",
        axis.line = element_line(color = "#4d4d4d", size = 1),
        axis.text.x = element_text(color = "#4d4d4d", size = 12),
        axis.text.y = element_text(color = "#4d4d4d", size = 12),
        axis.title.x = element_text(color = "#4d4d4d", size = 14, 
                                    face = "bold"),
        axis.title.y = element_text(color = "#4d4d4d", size = 14, 
                                    face = "bold"),
        axis.ticks.x = element_line(color = "#4d4d4d", size = 1),
        axis.ticks.y = element_line(color = "#4d4d4d", size = 1)) + 
  scale_color_viridis_d() +
  stat_ellipse(aes(group = interaction(group), color = group, fill = group), 
               alpha = 0.5, size = 1.1, linetype = 1, level = 0.40, 
               type = "t", geom = "polygon") +
  stat_ellipse(aes(group = interaction(group), color = group, fill = group), 
               alpha = 0.25, size = 0.75, linetype = 3, level = 0.95, 
               type = "t", geom = "polygon") + 
  geom_text(data = sand_label_df,aes(x = d13C, y = d15N, 
                                label = label, color = group, hjust = hjust,
                                vjust = vjust),
            size = 9/.pt, fontface = "bold") +
  scale_fill_viridis_d() +
  scale_x_continuous(limits=c(-27.5, -4), 
                     breaks = c(-25, -20, -15, -10, -5)) +
  scale_y_continuous(limits=c(-3, 15))

sand_p2
```

## Hummingbird Pueblo Figures
```{r, fig.width = 7, fig.height = 5}
hum_label_df <- data.frame(
  group = c("Hummingbird Pueblo Lepus", "Hummingbird Pueblo Sylvilagus",
            "Free-ranging Turkey", "Humans", "Maize-fed Turkey", "Bean",
            "Squash", "Corn"),
  label = c("Hummingbird Pueblo\nJackrabbbits",
            "Hummingbird Pueblo\nCottontails",
            "Free-Ranging\nTurkeys", "Humans", "Maize-Fed Turkeys", "Beans",
            "Squash", "Maize"),
  d13C = c(-20, -14.25, -20, -5.8, -8, -25.25, -24, -9.75),
  d15N = c(1.5, -1.25, 10, 12.2, 4.1, -2, 13.25, -1),
  hjust = c(0.5, 0.5, 0.5, 0, 0, 1, 1, 1),
  vjust = c(1, 0, 0, 1, 1, 1, 0, 1))

hum_label_df$group <- factor(hum_label_df$group, 
                          levels = c("Hummingbird Pueblo Lepus",
                                     "Hummingbird Pueblo Sylvilagus",
                                     "Free-ranging Turkey", "Humans", 
                                     "Maize-fed Turkey", "Bean", "Squash",
                                     "Corn"))

hum_plot <- SIBER_data %>%
  filter(group %in% c("Hummingbird Pueblo Lepus", 
                      "Hummingbird Pueblo Sylvilagus",
                      "Bean", "Corn", "Squash", "Free-ranging Turkey",
                      "Maize-fed Turkey", "Humans"))
  
hum_plot$group <- factor(hum_plot$group, 
                          levels = c("Hummingbird Pueblo Lepus",
                                     "Hummingbird Pueblo Sylvilagus",
                                     "Free-ranging Turkey", "Humans", 
                                     "Maize-fed Turkey", "Bean", "Squash",
                                     "Corn")) 
  
hum_p1 <- ggplot(hum_plot, aes(x = d13C, y = d15N)) +
  geom_point(aes(fill = group, color = group), stroke = 1, size = 4, 
             alpha = 0.5, shape = 21) +
  geom_point(aes(color = group), fill = NA, stroke = 1, size = 4, 
             shape = 21) +
  ylab(expression(paste(delta^{15}, "N (\u2030)"))) +
  xlab(expression(paste(delta^{13}, "C (\u2030)"))) + 
  theme_classic() +
  theme(legend.position = "none",
        axis.line = element_line(color = "#4d4d4d", size = 1),
        axis.text.x = element_text(color = "#4d4d4d", size = 12),
        axis.text.y = element_text(color = "#4d4d4d", size = 12),
        axis.title.x = element_text(color = "#4d4d4d", size = 14, 
                                    face = "bold"),
        axis.title.y = element_text(color = "#4d4d4d", size = 14, 
                                    face = "bold"),
        axis.ticks.x = element_line(color = "#4d4d4d", size = 1),
        axis.ticks.y = element_line(color = "#4d4d4d", size = 1)) + 
  scale_color_viridis_d() +
  stat_ellipse(aes(group = interaction(group), color = group, fill = group), 
               alpha = 0.25, size = 0.75, linetype = 3, level = 0.95, 
               type = "t", geom = "polygon") + 
  geom_text(data = hum_label_df,aes(x = d13C, y = d15N, 
                                label = label, color = group, hjust = hjust,
                                vjust = vjust),
            size = 9/.pt, fontface = "bold") +
  scale_fill_viridis_d() +
  scale_x_continuous(limits=c(-27.5, -4), 
                     breaks = c(-25, -20, -15, -10, -5)) +
  scale_y_continuous(limits=c(-3, 15))

hum_p1

hum_p2 <- ggplot(hum_plot, aes(x = d13C, y = d15N)) +
  ylab(expression(paste(delta^{15}, "N (\u2030)"))) +
  xlab(expression(paste(delta^{13}, "C (\u2030)"))) + 
  theme_classic() +
  theme(legend.position = "none",
        axis.line = element_line(color = "#4d4d4d", size = 1),
        axis.text.x = element_text(color = "#4d4d4d", size = 12),
        axis.text.y = element_text(color = "#4d4d4d", size = 12),
        axis.title.x = element_text(color = "#4d4d4d", size = 14, 
                                    face = "bold"),
        axis.title.y = element_text(color = "#4d4d4d", size = 14, 
                                    face = "bold"),
        axis.ticks.x = element_line(color = "#4d4d4d", size = 1),
        axis.ticks.y = element_line(color = "#4d4d4d", size = 1)) + 
  scale_color_viridis_d() +
  stat_ellipse(aes(group = interaction(group), color = group, fill = group), 
               alpha = 0.5, size = 1.1, linetype = 1, level = 0.40, type = "t", 
               geom = "polygon") +
  stat_ellipse(aes(group = interaction(group), color = group, fill = group), 
               alpha = 0.25, size = 0.75, linetype = 3, level = 0.95, 
               type = "t", geom = "polygon") + 
  geom_text(data = hum_label_df,aes(x = d13C, y = d15N, 
                                label = label, color = group, hjust = hjust,
                                vjust = vjust),
            size = 9/.pt, fontface = "bold") +
  scale_fill_viridis_d() +
  scale_x_continuous(limits=c(-27.5, -4), 
                     breaks = c(-25, -20, -15, -10, -5)) +
  scale_y_continuous(limits=c(-3, 15))

hum_p2
```

## Tijeras Pueblo Figures
```{r, fig.width = 7, fig.height = 5}
tij_label_df <- data.frame(
  group = c("Tijeras Pueblo Lepus", "Tijeras Pueblo Sylvilagus",
            "Free-ranging Turkey", "Humans", "Maize-fed Turkey", "Bean",
            "Squash", "Corn"),
  label = c("Tijeras Pueblo\nJackrabbbits",
            "Tijeras Pueblo\nCottontails",
            "Free-Ranging\nTurkeys", "Humans", "Maize-Fed Turkeys", "Beans",
            "Squash", "Maize"),
  d13C = c(-21, -17.5, -17.25, -5.8, -8, -25.25, -24, -9.75),
  d15N = c(1.5, -1.75, 9.5, 12.2, 4.1, -2, 13.25, -1),
  hjust = c(0.5, 0.5, 0.5, 0, 0, 1, 1, 1),
  vjust = c(1, 0, 0, 1, 1, 1, 0, 1))

tij_label_df$group <- factor(tij_label_df$group, 
                          levels = c("Tijeras Pueblo Lepus",
                                     "Tijeras Pueblo Sylvilagus",
                                     "Free-ranging Turkey", "Humans", 
                                     "Maize-fed Turkey", "Bean", "Squash",
                                     "Corn"))

tij_plot <- SIBER_data %>%
  filter(group %in% c("Tijeras Pueblo Lepus", 
                      "Tijeras Pueblo Sylvilagus",
                      "Bean", "Corn", "Squash", "Free-ranging Turkey",
                      "Maize-fed Turkey", "Humans"))
  
tij_plot$group <- factor(tij_plot$group, 
                          levels = c("Tijeras Pueblo Lepus",
                                     "Tijeras Pueblo Sylvilagus",
                                     "Free-ranging Turkey", "Humans", 
                                     "Maize-fed Turkey", "Bean", "Squash",
                                     "Corn")) 
  
tij_p1 <- ggplot(tij_plot, aes(x = d13C, y = d15N)) +
  geom_point(aes(fill = group, color = group), stroke = 1, size = 4, 
             alpha = 0.5, shape = 21) +
  geom_point(aes(color = group), fill = NA, stroke = 1, size = 4, 
             shape = 21) +
  ylab(expression(paste(delta^{15}, "N (\u2030)"))) +
  xlab(expression(paste(delta^{13}, "C (\u2030)"))) + 
  theme_classic() +
  theme(legend.position = "none",
        axis.line = element_line(color = "#4d4d4d", size = 1),
        axis.text.x = element_text(color = "#4d4d4d", size = 12),
        axis.text.y = element_text(color = "#4d4d4d", size = 12),
        axis.title.x = element_text(color = "#4d4d4d", size = 14, 
                                    face = "bold"),
        axis.title.y = element_text(color = "#4d4d4d", size = 14, 
                                    face = "bold"),
        axis.ticks.x = element_line(color = "#4d4d4d", size = 1),
        axis.ticks.y = element_line(color = "#4d4d4d", size = 1)) + 
  scale_color_viridis_d() +
  stat_ellipse(aes(group = interaction(group), color = group, fill = group), 
               alpha = 0.25, size = 0.75, linetype = 3, level = 0.95, 
               type = "t", geom = "polygon") + 
  geom_text(data = tij_label_df,aes(x = d13C, y = d15N, 
                                label = label, color = group, hjust = hjust,
                                vjust = vjust),
            size = 9/.pt, fontface = "bold") +
  scale_fill_viridis_d() +
  scale_x_continuous(limits=c(-27.5, -4), 
                     breaks = c(-25, -20, -15, -10, -5)) +
  scale_y_continuous(limits=c(-3, 15))

tij_p1

tij_p2 <- ggplot(tij_plot, aes(x = d13C, y = d15N)) +
  ylab(expression(paste(delta^{15}, "N (\u2030)"))) +
  xlab(expression(paste(delta^{13}, "C (\u2030)"))) + 
  theme_classic() +
  theme(legend.position = "none",
        axis.line = element_line(color = "#4d4d4d", size = 1),
        axis.text.x = element_text(color = "#4d4d4d", size = 12),
        axis.text.y = element_text(color = "#4d4d4d", size = 12),
        axis.title.x = element_text(color = "#4d4d4d", size = 14, 
                                    face = "bold"),
        axis.title.y = element_text(color = "#4d4d4d", size = 14, 
                                    face = "bold"),
        axis.ticks.x = element_line(color = "#4d4d4d", size = 1),
        axis.ticks.y = element_line(color = "#4d4d4d", size = 1)) + 
  scale_color_viridis_d() +
  stat_ellipse(aes(group = interaction(group), color = group, fill = group), 
               alpha = 0.5, size = 1.1, linetype = 1, level = 0.40, 
               type = "t", geom = "polygon") +
  stat_ellipse(aes(group = interaction(group), color = group, fill = group), 
               alpha = 0.25, size = 0.75, linetype = 3, level = 0.95, 
               type = "t", geom = "polygon") + 
  geom_text(data = tij_label_df,aes(x = d13C, y = d15N, 
                                label = label, color = group, hjust = hjust,
                                vjust = vjust),
            size = 9/.pt, fontface = "bold") +
  scale_fill_viridis_d() +
  scale_x_continuous(limits=c(-27.5, -4), 
                     breaks = c(-25, -20, -15, -10, -5)) +
  scale_y_continuous(limits=c(-3, 15))

tij_p2
```

## Combine Figures
```{r, fig.width = 14, fig.height = 15}
all_plots <- plot_grid(sand_p1, sand_p2, hum_p1, hum_p2, tij_p1, tij_p2,
                       labels = "AUTO", label_colour = "#4d4d4d", 
                       label_size = 20, ncol = 2, nrow = 3)

all_plots

ggsave("Figure 2.jpg", dpi = 300)
```

## Test for Normality
Some isotope values per group are non-normal. Thus, ellipses in
Figure 2 are visualized based on the t-distribution, which is also good for
small sample sizes.
```{r}
SIBER_data %>%
  group_by(group) %>%
  do(tidy(shapiro.test(.$d13C)))

SIBER_data %>%
  group_by(group) %>%
  do(tidy(shapiro.test(.$d15N)))
```

# SIBER Area Calculations
TA = Total Area, SEA = Standard Ellipse Area, and SEAc = Small Sample Size
Corrected Standard Ellipse Area (see Jackson et al. 2011)
```{r, fig.height=8, fig.width=10}
siber.example <- SIBER_data %>%
  select(d13C, d15N, group) %>%
  mutate(community = 1) %>%
  rename(iso1 = d13C,
         iso2 = d15N)
  
siber.example <- createSiberObject(siber.example)

group.ML1 <- data.frame(groupMetricsML(siber.example)) %>%
  rename("Hummingbird Jackrabbits" = X1.Hummingbird.Pueblo.Lepus,
         "Hummingbird Cottontails" = X1.Hummingbird.Pueblo.Sylvilagus,
         "Sand Canyon Jackrabbit" = X1.Sand.Canyon.Pueblo.Lepus,
         "Sand Canyon Cottontails" = X1.Sand.Canyon.Pueblo.Sylvilagus,
         "Tijeras Pueblo Jackrabbits" = X1.Tijeras.Pueblo.Lepus,
         "Tijeras Pueblo Cottontails" = X1.Tijeras.Pueblo.Sylvilagus,
         "Beans" = X1.Bean,
         "Maize" = X1.Corn,
         "Squash" = X1.Squash,
         "Free-ranging Turkey" = X1.Free.ranging.Turkey,
         "Maize-fed Turkey" = X1.Maize.fed.Turkey,
         "Humans" = X1.Humans) %>%
  t() %>%
  round(digits = 2)
  
group.ML1
```

# SIBER Maximum Likelihood Overlap with Humans Calculations
```{r}
results <- data.frame()
taxa <- c("1.Hummingbird Pueblo Lepus", "1.Hummingbird Pueblo Sylvilagus", 
              "1.Sand Canyon Pueblo Lepus", 
              "1.Sand Canyon Pueblo Sylvilagus",
              "1.Tijeras Pueblo Lepus", 
              "1.Tijeras Pueblo Sylvilagus",
              "1.Maize-fed Turkey")

for (i in seq_along(taxa)) {
  
  sea.overlap <- maxLikOverlap(taxa[[i]], "1.Humans", siber.example, 
                             p.interval = 0.95, n = 100)
  
  results[i, 1] <- taxa[[i]]
  results[i, 2] <- round(sea.overlap[[3]], digits = 2)
  results[i, 3] <- round(sea.overlap[[3]]/sea.overlap[[2]]*100, digits = 2)
  results[i, 4] <- round(sea.overlap[[3]]/sea.overlap[[1]]*100, digits = 2)
  results[i, 5] <- round(sea.overlap[[3]]/(sea.overlap[[2]] + 
                                             sea.overlap[[1]] - 
                                             sea.overlap[[3]])*100, digits = 2)
}

colnames(results) <- c("group", "overlap ‰", "% human niche", "% group niche", 
                       "% overlap")

results$group <- gsub("1.","", as.character(results$group))
results$group <- gsub("Lepus","Jackrabbits", as.character(results$group))
results$group <- gsub("Sylvilagus","Cottontails", as.character(results$group))

results

round(mean(results$`% human niche`[1:6]), digits = 2)
round(mean(results$`% group niche`[1:6]), digits = 2)
```
# References
Ambrose SH (1990) Preparation and characterization of bone and tooth collagen
  for isotopic analysis. Journal of Archaeological Science 17(4):431-451.

Basinger MA & Robertson PA (1997) Vascular flora of an old-growth forest 
  remnant in the Ozark hills of southern Illinois-updated results. 
  Transactions of the Illinois State Academy of Science 90(1-2):1-20.

Bruhl JJ & Wilson KL (2007) Towards a comprehensive survey of C3 and C4 
  photosynthetic pathways in Cyperaceae. Aliso: A Journal of Systematic and 
  Evolutionary Botany 23(1):99-148.

Chisholm B & Matson RG (1994) Carbon and Nitrogen Isotopic Evidence on
  Basketmaker II Diet at Cedar Mesa, Utah. KIVA 60(2):239-255.

Coltrain JB, Janetski JC, & Carlyle SW (2007) The Stable- and Radio-Isotope
  Chemistry of Western Basketmaker Burials: Implications for Early Puebloan 
  Diets and Origins. American Antiquity 72(2):301-321.

Conrad C, Jones EL, Newsome SD, & Schwartz DW (2016) Bone isotopes, eggshell 
  and turkey husbandry at Arroyo Hondo Pueblo. Journal of Archaeological 
  Science:Reports 10:566-574.

Danneberger K (1999) Weeds: shedding light on an old foe. Turf Grass Trends 
  8(11):1-3.

Dombrosky J (2020) A ~1000-year 13C Suess correction model for the study of
  past ecosystems. The Holocene 30(3):474-478.

Giussani LM, Cota-Sánchez JH, Zuloaga FO, & Kellogg EA (2001) A molecular 
  phylogeny of the grass subfamily Panicoideae (Poaceae) shows multiple origins 
  of C4 photosynthesis. American Journal of Botany 88(11):1993-2012.

Jackson AL, Inger R, Parnell AC, & Bearhop S (2011) Comparing isotopic niche 
  widths among and within communities: SIBER – Stable Isotope Bayesian Ellipses 
  in R. Journal of Animal Ecology 80(3):595-602.

Jones EL, Conrad C, Newsome SD, Kemp BM, & Kocer JM (2016) Turkeys on the 
  fringe: Variable husbandry in “marginal” areas of the prehistoric American 
  Southwest. Journal of Archaeological Science: Reports 10:575-583.

Kellner CM, Schoeninger M, Spielmann KA, & Moore K (2010) Stable isotope data 
  show temporal stability in diet at Pecos Pueblo and diet variation among 
  Southwest pueblos. Pecos Pueblo revisited: the biological and social context,
  ed Morgan ME (Peabody Museum Press, Cambridge, MA), Vol 85, pp 79-91.

Kennett DJ, et al. (2017) Archaeogenomic evidence reveals prehistoric 
  matrilineal dynasty. Nature Communications 8:14115.

Kocacinar F & Sage RF (2003) Photosynthetic pathway alters xylem structure and
  hydraulic function in herbaceous plants. Plant, Cell & Environment 
  26(12):2015-2026.

Martin SL (1999) Virgin Anasazi Diet as Demonstrated Through the Analysis of
  Stable Carbon and Nitrogen Isotopes. KIVA 64(4):495-514.

McCaffery H, Tykot RH, Gore KD, & DeBoer BR (2014) Stable Isotope Analysis of 
  Turkey (Meleagris gallopavo) Diet from Pueblo II and Pueblo III Sites, Middle
  San Juan Region, Northwest New Mexico. American Antiquity 79(2):337-352.

Nelson DM (2012) Carbon isotopic composition of Ambrosia and Artemisia pollen:
  assessment of a C3-plant paleophysiological indicator. New Phytologist 
  195(4):787-793.

Osborne CP, et al. (2014) A global database of C4 photosynthesis in grasses. 
  New Phytologist 204(3):441-446.

Rawlings TA & Driver JC (2010) Paleodiet of domestic turkey, Shields Pueblo 
  (5MT3807), Colorado: isotopic analysis and its implications for care of a 
  household domesticate. Journal of Archaeological Science 37(10):2433-2441.

Scribner KT & Krysl LJ (1982) Summer Foods of the Aububons Cottontail 
  (Sylvilagus auduboni: Leporidae) on Texas Panhandle Playa Basins. 
  The Southwestern Naturalist 27(4):460-463.

Syvertsen JP, Nickell GL, Spellenberg RW, & Cunningham GL (1976) Carbon 
  Reduction Pathways and Standing Crop in Three Chihuahuan Desert Plant 
  Communities. The Southwestern Naturalist 21(3):311-320.
